\documentclass{article}
\usepackage{geometry}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\graphicspath{ {images/} }
\geometry{a4paper, left=25mm, right=25mm, top=25mm, bottom=25mm}
\setlength{\parindent}{2em}
\setlength{\parskip}{1em}
\renewcommand{\baselinestretch}{1.3}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\rhead{S.Goldie - 42611814}
\lhead{FOAR705 - Learning Journal}
\lfoot{Session 2, 2019 - Macquarie University}
\rfoot{Page \thepage}


\title{FOAR705 - Digital Fundamentals}
\author{Sheriden Goldie}
\date{Session 2, Macquarie University}

\begin{document}

\maketitle

\section*{Data Carpentry Exercises}
\tableofcontents
\pagebreak

\section{Spreadsheets for Social Sciences}

It is important for researchers to observe good data practices when working with data to ensure that errors, corruption, and misrepresentation are kept to a minimum. 
Good data practices mean that if mistakes are made - they can be found, noticed, and rectified before any analysis is done. This saves time and frustration from the researchers, and those that hope to use their method or findings. 

\textbf{Initial Questions}
\begin{itemize}
    \item How many people have used spreadsheets in their research?
\end{itemize}

I haven't used spreadsheets for my research - but in my previous 'life' i worked in inventory planning and operations. This lead me to using spreadsheets everyday.

\begin{itemize}
    \item How many people have accidentally done something that made them frustrated or sad?
\end{itemize}

I have had the experience of someone messing up my model data for inventory sheets, that was an time-sink to find the issues and to fix. I have also accidentally applied formulas to the wrong cells, which gave me really weird results. 

\subsection{Formatting data tables in Spreadsheets}

\textbf{Responses to Questions}

\begin{enumerate}
    \item With the person next to you, identify what is wrong with this spreadsheet. Discuss the steps you would need to take to clean up the two tabs, and to put them all together in one spreadsheet.
\end{enumerate} 

There are many things wrong with this spreadsheet:
\begin{itemize}
    \item It is difficult that all the data is split across different tables in each tab - it is hard to tell if the studied locations overlap or not. 
    \item Terms in columns vary - eg. \begin{verbatim} mabati_sloping vs. mabatisloping
    \end{verbatim}
    \item Some data doesn't make sense - eg. -99 in the rooms column - this is not a full integer and is not a valid data point. 
    item\ Data tables are not the same in both tabs. There is a 'Plots' table in Mozambique tab, but not in the Tanzania tab.
    \item the highlighted cell is unclear as to the relevance of the additional barn
    \item the asterisk marked data to include a cowshed is unclear
    \item the asterisked data on cows is misleading as it adds the now dead cow to the count that should just be looking at the collected temporally fixed data
    \item the livestock table in the Mozambique tab is VERY problematic. It combines all data in one cell without defining the individual animal counts
    \item the Mozambique livestock table also includes the 'Look after Cows' column which is not consistent/relevant
    \item the livestock table in the Tanzania tab contains empty fields - is this a zero count, or a missing count?
    \item In the Tanzania tab there seems to be an inconsistency with the 'sunbricks' entry - is this different to burntbricks? - how are these forms being classified and named for the purpose of the study?
    \end{itemize}

\subsubsection{Metadata}
\textbf{A Note on Metadata}

Metadata is data about your data. This is the information that is going to tell you what the answer to a query that is in a column is in reply to. This is imperative as human memory is fallible, and you are likely to forget your intentions unless they are noted and recorded explicitly. 

Digital data is designed to be read by machines, but understanding the meaning of the data can only be done by humans, and this is the valuable part of the process. However your data should be understandable to other humans - for the reasons that they might need to try and replicate your results, or they may wish to replicate the model of your study, or even for their own project starting point - being able to understand the data is made more efficient with metadata. 

Considering metadata during both the collection and analysis phases of research is important, especially if your research is to become part of the scholarly record. 

\noindent \textbf{Questions}
\begin{itemize}
    \item Discuss this data with a partner and make a list of some of the types of metadata that should be recorded about this data set. It may be helpful to start by asking yourself, ``What is not immediately obvious to me about this data? What questions would I need to know the answers to in order to analyze and interpret this data?''
\end{itemize}

Types of metadata that should be recorded
\begin{itemize}
    \item The question or prompt that generated each type of response - eg: for no. of members - the question might be: How many people live in this household? or it might be how many people are in your family? these are distinctly different questions
    \item what the expected range or allowable response is - eg: numerical only, between 1-30, yes/no response only
    \item what the terms used in the data mean - eg: muddaub - means the walls are structured using mud daubed over straw. Or definitions like what a 'plot' or 'room' is being defined as. 
\end{itemize}

\subsection{Formatting Problems}

Other problems to consider is the formatting - using the ``mergecell'' function to make the data look ``pretty''.
As this is ``raw data'' its purpose is to organise the data in a way that is effective and able to be manipulated by the program to create relevant, interpret able outputs. 

\subsection{Dates as Data}
\textbf{Separating dates into components}
to do the exercise these are the processes/actions I took:
\begin{enumerate}
    \item Download file
    \item Open file using Microsoft Excel
    \item Identified table to manipulate
    \item Copied table to new sheet - labelled this sheet 'working tab'
    \item Added three columns between columns A and B using insert columns function. These are labelled 'Day', 'Month', and 'Year' respectively.
    \item Entered the following codes into cells b2: =DAY(A2) c2: =MONTH(A2) d2: =YEAR(A2)
    \item I formatted columns B, C, and D, to a number format without decimal places. 
    \item Adding a new line of data just with the 17/11 information meant the data in the year column automatically populated the current year - 2019. 
\end{enumerate}

\subsection{Quality Assurance}
The processes/actions I took:
\begin{enumerate}
    \item Download file
    \item Open file using Microsoft Excel
    \item Copied table to new sheet - labelled this sheet 'working tab'
    \item Select column D
    \item While column is highlighted go to File - Data - Data Tools - Data Validation
    \item In the section tab of the window use the drop-down menu under Allow to select 'Whole Number'.
    \item Under Data select 'between' and enter 1 as the minimum and 30 as the maximum. 
    \item To test this is working, I tried entering '31' in cell D133. I received an error message as expected. 
    \item Go back to the the Data Validation window. In the Input Error tab, I enter a custom input error name and description.
    \item Go back to the the Data Validation window. In the Error Alert tab select style: Warning and input a custom warning. Title: Invalid Entry. Error Message: Entry must be between 1 and 30.
    \item I created an additional rule set for column E. The age range was set as 0 - 120, it should be a whole number and the warning message was set as: Age should be in whole years and between 0 and 120.
    \item I applied an additional data validation to column F. This was to allow a selection from a list of accepted entries. 
    \item I went to the Data Validation Window while the column was selected. In the Setting's tab I selected the following: ``Allow'' was set to ``List''. with ``ignore blank'' and ``in cell dropdown'' checked. 
    \item Then in the ``Source'' field, I typed the entries, separated by a comma and a space. As so "grass, muddaub, burntbricks, sunbricks, cement"
    \item I entered the Input Error as, title: ``Entry must be as per List'', and Error message: ''The entry must be entered as per the following list: grass, muddaub, burntbricks, sunbricks, cement. Use the dropdown menu to select your entry''
    \item I also tried creating a table as the  ``Source'' data in an additional worksheet. This worked - however care must be taken as this would not be preserved if exporting the data to a .CSV file. 
\end{enumerate}

\subsection{Exporting Data}
Save a file in CSV format.
The steps/actions I took:
\begin{enumerate}
    \item Go to: File - Save As
    \item Select folder directory to save file to
    \item Name file and select .CSV as file format.
    \item Click save
    \item As I had created a new tab for my working data I was only able to save that tab as a CSV. This is good to know for purposes of separating raw data, working data, and for versioning and back up purposes as well. 
\end{enumerate}

\subsection{Reflections}
The Data Carpentry exercises were useful as they reminded me of the 'best practice' ways of dealing with data, and the ideals for how it should be recorded to maximise the potential output. 
In my previous  roles in Inventory Planning and Finance Administration, I have worked with data a lot in various settings and for various purposes - so the mechanics of this task was not new or difficult for me, and so I did not encounter any major issues. 
However the process of documenting my learning process concurrent with the activity of learning was new, and it was useful to think through each step of the process as a single function. 

\section{The Unix Shell}

The aim of this exercise is to gain familiarity with using a Unix Shell. 
I have heard of a Unix Shell before, and have seen one used to manage the data for a commission based sales system. However I have never used it, and am interested to see how it actually works. 

\subsection{Setup}
In the setup we download a zip file, move it to the desktop and extract the files within it. 
The next step is to open a 'terminal' and type a command. 

I am using windows and do not have a native terminal - so I need to download one. The lesson recommends Git for Windows. 

\section{OpenRefine for Social Sciences}
\subsection{Part 1}


\section{R for Social Sciences}
\subsection{Part 1}


\end{document}
